TacticToe: Learning to Prove with Tactics

### 2018-08-20 tactictoe_jv.tex
Term instantiation

Here, the two examples accepts only very specific kind of theorems: the theorem 
argument should match the precondition of the goal for \texttt{MATCH\_MP\_TAC} 
and an induction principle should be given to \texttt{INDUCT\_THEN}.
The likelihood of finding such theorem is low, thus given the list of 
predictions $\mathit{Predictions}^{\mathit{theorem}}_{16}(g)= 
[thm1,\ldots,thm16]$, we create 16 instantiations and try them in the 
order given by the predictor, and we return the result of the first successful 
tactic and produces new goals. With a modified version of 
the tactical \texttt{ORELSE} that fails if the tactic 
did not have any effect, we
can write for the first example a single new tactic encoding this procedure: 
  \[\texttt{MATCH\_MP\_TAC thm1 ORELSE ... ORELSE MATCH\_MP\_TAC thm16}\]

\paragraph{Term}
Some proofs requires to instantiate a tactic argument by a term. Here are two 
abstracted 
tactics that uses term 
arguments with placeholders \texttt{term} and \texttt{thm}: 
  \[\texttt{EXISTS\_TAC term},\ \texttt{SPEC\_THEN term STRUCT\_CASES\_TAC 
  thm}\]
Finding a witness to instantiate an existential formula is difficult because it 
is likely that this witness may not have been proposed before and there is no 
general rule on how to generate it. Instantiating 
the term argument in the second tactic is easier because the expected term is a 
variable of the goal. That is why we think that the first step toward term 
prediction is to select subterms of the goal ordered by their similarity with 
the term argument appearing in the original tactic. 
Side experiments using a tactic that tries the 16 closest subterms 
$[term1,\ldots,term16]$ to the goal with the procedure described in the 
previous paragraph 
shows that this simple term predictor improves the success rate of 
\tactictoe by another 4 percent. The tactic derived from the abstraction
\texttt{EXISTS\_TAC term} in this experiment is:
   \[\texttt{EXISTS\_TAC term1 ORELSE ... ORELSE EXISTS\_TAC term16}\]

>>>>

Here, the two examples accepts only very specific kind of theorems: the theorem 
argument should match the precondition of the goal for \texttt{MATCH\_MP\_TAC} 
and an induction principle should be given to \texttt{INDUCT\_THEN}.
The likelihood of finding such theorem is low, thus given the list of 
predictions $\mathit{Predictions}^{\mathit{theorem}}_{16}(g)= 
[thm1,\ldots,thm16]$, we create 16 instantiations and try them in the 
order given by the predictor, and we return the result of the first successful 
tactic and produces new goals. With a modified version of 
the tactical \texttt{ORELSE} that fails if the tactic 
did not have any effect, we
can write for the first example a single new tactic that implements this 
procedure: 
  \[\texttt{MATCH\_MP\_TAC thm1 ORELSE ... ORELSE MATCH\_MP\_TAC thm16}\]

\paragraph{Term}
Some proofs requires to instantiate a tactic argument by a term. Here are two 
abstracted 
tactics that uses term 
arguments with placeholders \texttt{term} and \texttt{thm}: 
  \[\texttt{EXISTS\_TAC term},\ \texttt{SPEC\_THEN term STRUCT\_CASES\_TAC 
  thm}\]
Finding a witness to instantiate an existential formula is challenging because 
it is likely that this witness may not have been proposed before and there is 
no 
general rule on how to generate it. Instantiating 
the term argument in the second tactic is easier because the expected term is a 
variable of the goal. That is why we think that the first step 
toward term argument prediction is to select subterms of the goal ordered by 
their similarity with the term argument appearing in the original tactic. 
Side experiments, using a tactic that tries the 16 closest subterms 
$[term1,\ldots,term16]$ to the goal with the trial and error procedure 
described in the previous paragraph, shows that this simple term predictor 
improves the success 
rate of 
\tactictoe by another four percent. With the modified 
\texttt{ORELSE} tactical, the tactic derived from the abstraction
\texttt{EXISTS\_TAC term} in this experiment can be written as:
   \[\texttt{EXISTS\_TAC term1 ORELSE ... ORELSE EXISTS\_TAC term16}\]
